\section{Methodology} \label{sec:methodology}
This research employs an experimental approach to evaluate the effectiveness of VxLang's code virtualization. We compare the reverse engineering difficulty and performance characteristics of software binaries before and after applying VxLang's virtualization.

\subsection{Experimental Design}
A comparative study design was used, involving a control group (original, non-virtualized binaries) and an experimental group (binaries with critical sections virtualized by VxLang).

\begin{itemize}
    \item \textbf{Independent Variable:} Application of VxLang code virtualization (Applied vs. Not Applied).
    \item \textbf{Dependent Variables:}
        \begin{itemize}
            \item \textit{Reverse Engineering Protection Effectiveness:} Evaluated quantitatively and through systematic observation. This includes recording the success/failure rates of authentication bypass attempts, analyzing the behavior of static analysis tools (e.g., Ghidra's ability to disassemble code, identify critical strings, and map control flow), and observing outcomes of dynamic analysis (e.g., x64dbg's effectiveness in tracing execution, identifying runtime data, and success/failure of runtime manipulation). Metrics such as the number of obfuscated critical strings, tool-reported disassembly failures, and binary indicators of bypass success will be used.
            \item \textit{Performance Overhead:} Quantitatively measured via execution time for specific computational tasks (QuickSort, AES encryption/decryption).
            \item \textit{File Size Overhead:} Quantitatively measured by comparing the size (in bytes) of the final executable files.
        \end{itemize}
\end{itemize}

\subsection{Study Objects}
Two categories of applications were developed and analyzed:

\subsubsection{Authentication Case Study Applications} Simple applications simulating user login were created to serve as targets for reverse engineering analysis focused on bypassing the authentication mechanism. Variants included:
    \begin{itemize}
        \item \textbf{Interface Types:} Console (CLI), Qt Widgets (GUI), Dear ImGui (Immediate Mode GUI).
        \item \textbf{Authentication Mechanisms:} Hardcoded credentials (comparing input against string literals) and Cloud-based validation (sending credentials via HTTP POST to a local backend server).
    \end{itemize}
    For each variant, the core authentication logic (comparison function or the call to the cloud request function and subsequent result check) was targeted for virtualization in the experimental group.

\subsubsection{Performance Benchmark Applications} Applications designed to measure the performance impact of virtualization on specific computational tasks:
    \begin{itemize}
        \item \textbf{QuickSort Benchmark:} Implemented a standard recursive QuickSort algorithm. The core recursive function was virtualized. Tested with varying array sizes (100 to 1,000,000 elements).
        \item \textbf{AES Encryption Benchmark:} Implemented AES-256-CBC encryption/decryption using OpenSSL's EVP API. The loop performing batch encryption/decryption operations on 1GB of data was virtualized.
        \item \textbf{File Size Benchmark:} A minimal application with embedded dummy data to assess the baseline size increase due to the inclusion of the VxLang runtime.
    \end{itemize}

\subsubsection{Case Study: Lilith RAT}
To further evaluate VxLang's effectiveness on more complex software potentially exhibiting malicious characteristics and to assess its impact on automated detection tools, a Remote Administration Tool (RAT) named Lilith \cite{LilithRAT} was included as an additional study object. The client component of this open-source C++ RAT was compiled and analyzed both in its original form and after applying VxLang virtualization to its core functions. Analysis focused on static/dynamic reverse engineering difficulty, functional integrity post-virtualization, and detection rates by antivirus engines via VirusTotal. The Lilith server component remained unmodified and was used for functional testing of the client.

\subsection{Instrumentation and Materials}
\begin{itemize}
    \item \textbf{Hardware:} Standard Windows 11 (64-bit) PC.
    \item \textbf{Development Tools:} Clang/clang-cl (C++17), CMake, Ninja, Neovim.
    \item \textbf{Libraries/Frameworks:} VxLang SDK, Qt 6, Dear ImGui (+GLFW/OpenGL3 backend), OpenSSL 3.x, libcurl, nlohmann/json.
    \item \textbf{Analysis Tools:} Ghidra (v11.x) for static analysis, x64dbg (latest snapshot) for dynamic analysis.
    \item \textbf{Performance Measurement:} C++ \texttt{std::chrono::high\_resolution\_clock} for timing, \texttt{std::filesystem::file\_size} for file size.
\end{itemize}


\subsection{Data Collection Procedure}

\subsubsection{Security Analysis}
For each authentication application (original and virtualized):
    \begin{enumerate}
        \item \textbf{Static Analysis (Ghidra):} Load executable, search for relevant strings (e.g., "Failed", "Authorized", potential credentials), analyze disassembly/decompilation around string references or entry points, identify conditional jumps controlling authentication success/failure, attempt static patching to bypass logic. Record systematic observations regarding tool behavior (e.g., string search success, disassembly quality, logic identifiability) and the success/failure of static patching attempts.
        \item \textbf{Dynamic Analysis (x64dbg):} Run executable under debugger, search for strings/patterns at runtime, set breakpoints at suspected logic locations (identified via static analysis or runtime observation), step through execution, observe register/memory values, attempt runtime manipulation (patching conditional jumps) to bypass authentication. Record systematic observations regarding runtime behavior, data visibility (especially for virtualized strings), the success/failure of runtime manipulation for bypass, and other objective indicators of analysis complexity.
    \end{enumerate}

\subsubsection{Performance Analysis}
For each benchmark application (original and virtualized):
    \begin{enumerate}
        \item \textbf{Execution Time:} Run QuickSort benchmark 100 times per data size, record individual times. Run AES benchmark on 1GB data, record total encryption/decryption time. Use \texttt{std::chrono}. Calculate average, standard deviation (for QuickSort), and throughput (for AES).
        \item \textbf{File Size:} Measure the size of the final executable file in bytes using \texttt{std::filesystem::file\_size}.
    \end{enumerate}

\subsubsection{Lilith RAT Analysis}
\label{subsubsec:methodology_lilith_rat_journal} 
For the Lilith RAT client (original and virtualized):
\begin{enumerate}
    \item \textbf{Functional Integrity Testing:} The virtualized client was tested for core RAT functionalities (connection to server, remote command execution, file system access) against an unmodified server on a local network (client IP: \texttt{192.168.1.15}, server IP: \texttt{192.168.1.235} on port \texttt{1337}) to ensure virtualization did not break essential operations. A test file (\texttt{password.txt} containing \texttt{"THIS IS A SECRET"}) on the client machine was used to verify remote file access.
    \item \textbf{VirusTotal Analysis:} Both original and virtualized client executables were submitted to VirusTotal to compare detection rates and threat characterizations by various antivirus engines.
\end{enumerate}

\subsection{Data Analysis Techniques}
\begin{itemize}
    \item \textbf{Security Protection Analysis:} Quantitative metrics (e.g., bypass success rates, percentage of critical strings successfully obfuscated, reduction in identifiable functions) and descriptive analysis of systematically recorded observations from static and dynamic analysis (e.g., tool disassembly success, string visibility) will be used to compare the protection effectiveness between control and experimental groups.
    \item \textbf{Performance Data:} Calculation of descriptive statistics (mean, standard deviation), percentage overhead for execution time, throughput calculation (MB/s), and percentage increase in file size. Comparative tables and graphs will be used for presentation.
    \item \textbf{Trade-off Analysis:} Synthesis of security findings and performance results to evaluate the balance between protection enhancement and performance/size costs introduced by VxLang.
\end{itemize}

